The importance of measurement uncertainty in ecological management

### 2018-01-06 manuscripts/manuscript.Rmd

bibliography: citations.bib
output: rticles::elsevier_article
layout: 3p # review # review = doublespace, 3p = singlespace, 5p = two-column

@@

Fisheries conservation and management has long been both a crucible and proving ground for the theory of ecological management more generally, including topics such as adaptive management [@Walters1978], ecosystem-based management [@Levin2008] and resilience thinking [@Holling1973; @May1977], while also giving rise to the sub-discipline of resource economics [@Beverton1957; @Schaefer1957], thanks to its global relevance, long history, and readily available data.  Consequently, understanding the impact of measurement uncertainty in this rich and well-studied context will have implications for both resource management and conservation efforts in other domains, many of which draw on concepts established or tested in fisheries ecosystems.  In particular, the approach here illustrates the limitations of both over-simple optimal control approaches and rule-of-thumb precautionary approaches, and offers instead a more general strategy for dealing with measurement uncertainty adaptively.

@@

## Model

@@

## The decision problem

We seek to determine the sequence of actions $h_t$ for $t \in [1, ... \infty]$ that maximize the net present value (discounted sum of all future profits) of the fishery.  We will denote the discount rate $\delta$. For simplicity we will once again follow classic theory and assume a fixed price for fish, (equivalently, measuring our value in units of discounted fish rather than discounted dollars), though again this assumption is easy to relax. Each year, the manager also obtains an estimate $y_t$ of the true $x_t$ stock size,

@@

$$\max_{ \lbrace h_t \rbrace} \sum_{t=0}^{\infty} \delta x_t $$

@@

## Current theory and current practice

@@

## Pardoxes of uncertainty

@@

### Reed's Paradox: $S = D$

@@

### Clark's Paradox: $M < D$

@@

The resulting solutions only deepened the paradox, rather than recovering a more cautious policy in response to the uncertainty, @Clark1986 finds policies that are often *less* cautious:

@@

Using different but still implicitly approximate results nearly two decades later, @Sethi2005 largely confirm Clark's Paradox, which they likewise observe with some concern:

@@

While other attempts to introduce measurement error have been considered, each relies on various approximations that similarly alter the problem, such as assuming the solution must be of 'constant-escapement' type [@Ludwig1981; @Roughgarden1996; @Engen1997;  @Moxnes2003].  Here we will illustrate the use of powerful modern algorithms that will permit a more direct solution to the case of measurement uncertainty.  First we will introduce the more heuristic alternative to addressing uncertainty, which more accurately reflects both intuition and actual management.

## Uncertainty in practice

@@

Numerically, each of these functions are defined over a discrete set of possible states, observations, and actions, and can thus be represented as a collection of matrices or tensors.  As in the case of MDP [@Marescot2013], the POMDP problem can be solved by existing and well-studied algorithms. While not unknown to the conservation literature [@Williams2011], algorithms for POMDP have historically scaled quite poorly, and their application has been restricted to contexts with only a handful of possible states and actions [e.g. @Chades2008; @Chades2011; @Fackler2014; @Fackler2014b].  POMDP algorithms remain a very active area in artificial intellegence, and by adapting recent algorithmic developments [@Kurniawati2008] from that field we are able to find solutions for the considerably more complex problems such as the case of fisheries managment, which requires on the order of 100 states and actions to provide sufficient numerical resolution. Readers interested in the technical details of these algorithms should consult the extensive literature in in POMDP methods. For those interested in exploring or applying these methods, we have developed a preliminary R package, @sarsop-pkg, which can be used to replicate any of the results presented here and explore other variations.

## Results

@@

### Replicate simulations

@@

### Comparisons to historical data

@@

## Discussion

@@

### A practical implementation?

>>>>

bibliography: citations.bib
csl: the-american-naturalist.csl
output: rticles::elsevier_article
layout: 3p # review # review = doublespace, 3p = singlespace, 5p = two-column

@@

Fisheries conservation and management has long been both a crucible and proving ground for the theory of ecological management more generally, including topics such as adaptive management [@Walters1978], ecosystem-based management [@Levin2008] and resilience thinking [@Holling1973; @May1977], while also giving rise to the sub-discipline of resource economics [@Beverton1957; @Schaefer1954; @Gordon1954], thanks to its global relevance, long history, and readily available data.  Consequently, understanding the impact of measurement uncertainty in this rich and well-studied context will have implications for both resource management and conservation efforts in other domains, many of which draw on concepts established or tested in fisheries ecosystems.  In particular, the approach here illustrates the limitations of both over-simple optimal control approaches and rule-of-thumb precautionary approaches, and offers instead a more general strategy for dealing with measurement uncertainty adaptively.

@@

# Model

@@

# The Decision Problem

We seek to determine the sequence of actions $h_t$ for $t \in [1, ... \infty]$ that maximize the net present value (discounted sum of all future profits) of the fishery.  We will denote the discount rate $\gamma$. For simplicity we will once again follow classic theory and assume a fixed price for fish, (equivalently, measuring our value in units of discounted fish rather than discounted dollars), though again this assumption is easy to relax. Each year, the manager also obtains an estimate $y_t$ of the true $x_t$ stock size,

@@

$$\max_{ \lbrace h_t \rbrace} \mathbb{E} \left \{ \sum_{t=0}^{\infty} \gamma^t \cdot h_t \right \}$$

@@

<!-- MDP, POMDP -->

# Current theory and practice

@@

@@

## Reed's Paradox: $S = D$

@@

## Clark's Paradox: $M < D$

@@

By leveraging algorithms (and also computational hardware) developed over two decades later are we able to crack this mathematical difficulty Clark refers to directly without this "simplyfying assumption."  Unforunately for @Clark1986, this "simplifying assumption" they made instead only deepened the paradox, leading them to find policies that become *less* cautious as uncertainty increases:

@@

Using a different but still flawed assumption nearly two decades later, @Sethi2005 largely confirm Clark's Paradox, which they likewise observed with some concern:

@@

While other attempts to introduce measurement error have been considered, each relies on various approximations that similarly alter the problem, such as assuming the solution must be of 'constant-escapement' type [@Ludwig1981; @Roughgarden1996; @Engen1997;  @Moxnes2003].  Here we will illustrate the use of powerful modern algorithms that will permit a more direct solution to the case of measurement uncertainty.

First we will introduce the more heuristic alternative to addressing uncertainty, which more accurately reflects both intuition and actual management.

## Uncertainty in Practice

@@

Numerically, each of these functions are defined over a discrete set of possible states, observations, and actions, and can thus be represented as a collection of matrices or tensors.  As in the case of MDP [@Marescot2013], the POMDP problem can be solved by existing and well-studied algorithms. While not unknown to the conservation literature [@Williams2011], algorithms for POMDP have historically scaled quite poorly, and their application has been restricted to contexts with only a handful of possible states and actions [e.g. @Chades2008; @Chades2011; @Fackler2014; @Fackler2014b].  POMDP algorithms remain a very active area in artificial intellegence, and by adapting recent algorithmic developments [@Kurniawati2008] from that field we are able to find solutions for the considerably more complex problems such as the case of fisheries managment, which requires on the order of 100 states and actions to provide sufficient numerical resolution. Readers interested in the technical details of these algorithms should consult the extensive literature in in POMDP methods. For those interested in exploring or applying these methods, we have developed a preliminary R package [@sarsop-pkg], which can be used to replicate any of the results presented here and explore other variations.

# Results

@@

@@

@@

# Discussion

@@

## A practical implementation?
